\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-04-20 paper/paper.tex
Updates to paper.

\subsection{Ganged Operations}

>>>>

\subsection{Ganged Operations}
Key value stores (such as Rocksdb) usually provide a batched write operation
that atomically updates a set of key-value pairs. A batched write is often a key
primitive for building more sophisticated operations such as
transactions. Batched writes require special handling in Cyclone due to the
requirement that we route a write request by hashing the key to select both the
control plane core (physical log) as well as KV store core. A batched write has
multiple keys and therefore we need to synchronize the progress of the operation
across multiple physical logs (and associated replication quorums) as well as KV
store cores. In effect the serial history contains a grouped set of operation on
multiple keys spanning multiple cores as a single atomic operation. All serial
histories across replicas therefore continue to map to the same equivalent
serial history modulo reordering of operations to different keys.

A simple solution to this problem is to use a distributed transaction, using
two-phase commit to couple the otherwise independent replication quorums and KV
store cores. However this method has a number of problems. First, it introduces
an additional network round trip that is normally not necessary for a key value
store with a single shared memory image holding all the keys. Second, handling
failures with distributed transactions would have cause an explosion in
complexity in Cyclone, something we wished to avoid in the interests of
practical deployment to independent end users who need to understand and 
maintain Cyclone as part of their KV store infrastructure.

Our solution instead is to use a technique we termed ganged operations, based on
the observation that in the failure free case, executing a batched write across
multiple control plane cores and KV store cores does not look very much
different from a synchronous shared memory barrier. The challenge is to
effectively manage failures in replication without causing participating cores
to stop making progress. The first step is to reconfigure RAFTs leader election
protocol to ensure that the leaders for all the physical logs are co-located on
the same host. Clients always dispatch batched writes to a fixed control plane
core (called the injection core) which is then responsible for serially
forwarding the requests to the participating control plane cores (for
corresponding physical logs). The injection core uses DPDKs packet cloning
primitives to avoid dataplane work while forwarding copies of the same batched
write packet to multiple control plane cores. The injection core also adds a
``nonce'' - a unique timestamp to the packet. In addition the injection core
adds a unique view number to the packet corresponding to the term numbers of all
the participating RAFT instances (read from shared memory of the co-located leaders).
We describe the remaining flow for ganged operations in terms of the pseudocode
in Figure~\ref{fig:ganged_ops}. We defer discussion of how we generate the nonce
to later in this section. We also assume that a unique barrier is allocated in
shared memory for each ganged operation. We discuss later how this is done
without using dynamic allocation.

In the absence of failures Figure~\ref{fig:ganged_ops} is straightforward. One
replication is complete on all participating physical logs, the corresponding KV
store cores execute {\tt event\_replicated\_ganged\_op}. The use a bitmask for
the barrier setting the appropriate bit for themselves. Once all KV store cores
have arrived at the barrier the operation is executed. A distinguished leader
core (we use use the minumum numbered participating KV store core) is
responsible for sending back the response to the client. Two types of failure
need to be considered. First, one of the non-leader cores could fail. The
remaning cores would never progress past the barrier. To solve this the leader
core montors the remaning cores for their current view - which is the RAFT term
of the last executed log entry. If the view has moved past the view in the
packet, replication failed on that physical log. The leader then sets the bit
for the core that will no longer participate in the barrier to aid progress (a
technique we borrowed from lock-free data structures~\cite{lock-free}). It also
sets a failed flag to signal failure to the other participating cores. The
second failure case is that the leader core might not receive the operation if
it fails during replication from corresponding physical log. This case is
detected in a similar manner by the remaining cores by monitoring the published
view from the leader core in shared memory.

\begin{figure*}
  \centering
\begin{verbatim}
    // Control plane core
    event_recv_ganged_op(packet)
    {
      Begin replication of packet
    }

    // KV store core
    event_replicated_ganged_op (packet, barrier)
    {
      if(leader KV core)
        atomic set bit me in barrier.mask         
        do
         for each KV store core in packet
          if core_public_data[core].view > packet.view
           barrier.failed = true 
           atomic set bit for core in barrier.mask
        while barrrer.mask != mask of participating KV cores
        if barrier.failed
         Ganged replication failed. Send retry to client. return.
        else       
         execute operation
         send response
      else
        wait until 
         core_public_data[leader core].view > packet.view
         OR core_public_data[me].barrier == mask of participating KV cores
        if core_public_data[leader core].view > packet.view
         Ganged replication failed. Send retry to client. return.
        else
         execute operation
    }
\end{verbatim}
\caption{Ganged Operation}
\label{fig:ganged_ops}
\end{figure*}
